---
title: "Descriptions of Route"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
1) General qualitative description of physical characteristics
2) Great circle speed, total avg speed using BSAM corrected locations, split the route into 4 parts:
     A) WAP to tip of South America (Cape horn -55.9833)
     B) tip of South America to peninsula de paracas(this is when the whales come in and start to hug the coast for the first time) -13.8667
    C) Peninsula de paracas to salinas (this is where we think they start breeding behavior -2.190181)
    D) salinas and above (breeding behavior)
For each of these sections, I will do average speeds.
3) Percentage of Routes in national waters

Take out 121208 and rerun things
First, Great circle speeds. We can keep these just have to fix the one whale. 121207
```{r}
library(argosfilter)
library(tidyverse)
library(lubridate)
library(knitr)
opts_chunk$set(echo=F,warning=F,message=F,fig.width = 11,fig.height = 5,cache=F)
library(boot)
library(raster)
library(tidyr)
library(ggplot2)
# MASS This fucks with dplyr, load later
library(ggmap)
library(dplyr)
library(chron)
library(gridExtra)
library(stringr)
library(R2jags)
library(maptools)
library(reshape2)
#New model to be run, flag turned off if just updating.
newModel<-T
register_google(key="<insertgooglekeyhere>")
adat<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/bsam models/bsamspeeds_migonly.csv")
adat<-adat%>%filter(id!=121208)
bdat<-split(adat,adat$id)
#then define what you want to do to each list, for instance apply an r package and add results as new column
fg<-lapply(bdat,function(x){
  y<-x %>% 
    mutate(total_distance = sum(distance_km), total_time=sum(Hours_elapsed))
  return(y)
})

gdat<-bind_rows(fg)



gdat<-gdat%>%group_by(id)%>%filter(row_number()==1 | row_number()==n())

hdat<-split(gdat,gdat$id)
#then define what you want to do to each list, for instance apply an r package and add results as new column
gh<-lapply(hdat,function(x){
  y<-x %>% 
    mutate(GC_distance = distanceTrack(lat, lon))
  return(y)
})

idat<-bind_rows(gh)

cdat<-idat%>%mutate(km_hr=total_distance/total_time)
cdat<-cdat%>%mutate(m_s=km_hr*5/18)
write.csv(cdat,"/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/summaryGCmig.csv")
```

Now find general average speeds
```{r}
edat<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/bsam models/bsamspeeds_migonly.csv")
edat<-edat%>%filter(id!=121208)

fdat<-edat%>%group_by(id)%>%mutate(avg_speed=mean(Speed_km_h))
gdat<-fdat%>%filter(row_number()==1| row_number()==n())
hdat<-gdat%>%mutate(avg_speed_m_s=avg_speed*5/18)
write.csv(hdat,"/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/summaryavgsp.csv")

```

Now do speeds by section
```{r}
idat<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/bsam models/model4-29-more iterations/bsamspeeds.csv")
idat<-idat%>%filter(id!=121208)

bottom<-idat%>%filter(lat<=-55.9833)
middle<-idat%>%filter(lat>=-55.9833 & lat<=-13.8667)
mid<-middle%>%group_by(id)%>%filter(row_number()==1 | row_number()==n())
top<-idat%>%filter(lat>=-13.8667 & lat<=-2.190181)
breeding<-idat%>%filter(lat>=-2.190181)

#Now we will do the average speeds for each section
bottom<-bottom%>%group_by(id)%>%mutate(avg_speed=mean(Speed_km_h))
middle<-middle%>%group_by(id)%>%mutate(avg_speed=mean(Speed_km_h))
top<-top%>%group_by(id)%>%mutate(avg_speed=mean(Speed_km_h))
breeding<-breeding%>%group_by(id)%>%mutate(avg_speed=mean(Speed_km_h))

#Now lets look at just first and last for each
bot<-bottom%>%group_by(id)%>%filter(row_number()==1 | row_number()==n())
bot<-bot%>%mutate(avg_speed_m_s=avg_speed*5/18)

mid<-middle%>%group_by(id)%>%filter(row_number()==1 | row_number()==n())
mid<-mid%>%mutate(avg_speed_m_s=avg_speed*5/18)
t<-top%>%group_by(id)%>%filter(row_number()==1 | row_number()==n())
t<-t%>%mutate(avg_speed_m_s=avg_speed*5/18)
breed<-breeding%>%group_by(id)%>%filter(row_number()==1 | row_number()==n())
breed<-breed%>%mutate(avg_speed_m_s=avg_speed*5/18)

#Now we should save them and then make a sheet. 
#write.csv(bot,"/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/avgspeeds_bottom.csv")
#write.csv(mid,"/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/avgspeeds_middle.csv")
#write.csv(t,"/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/avgspeeds_top.csv")
#write.csv(breed,"/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/avgspeeds_breeding.csv")

#Actually we are going to use a joining function for these
#first we have to rename
bot1<-bot%>%dplyr::select(3,19:20)
names(bot1) <- c("id", "Bottom_avg_sp_km_hr", "Bottom_avg_sp_m_s")

mid1<-mid%>%dplyr::select(3,19:20)
names(mid1) <- c("id", "Middle_avg_sp_km_hr", "Middle_avg_sp_m_s")

top1<-t%>%dplyr::select(3,19:20)
names(top1) <- c("id", "Top_avg_sp_km_hr", "Top_avg_sp_m_s")

breed1<-breed%>%dplyr::select(3,19:20)
names(breed1) <- c("id","Breeding_avg_sp_km_hr", "Breeding_avg_sp_m_s")


a<-bot1%>%
  left_join(mid1, c("id"="id"))
b<-a%>%left_join(top1,by="id")
speeds_section<-b%>%left_join(breed1, by="id")
speeds_section<-speeds_section%>%group_by(id)%>%filter(row_number()==1)
#rearrange order
speeds_section<-speeds_section%>%dplyr::select("id","Bottom_avg_sp_km_hr","Middle_avg_sp_km_hr","Top_avg_sp_km_hr","Breeding_avg_sp_km_hr","Bottom_avg_sp_m_s","Middle_avg_sp_m_s","Top_avg_sp_m_s","Breeding_avg_sp_m_s")


#oops we already got rid of the breeding area values
br<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/bsam models/bsamspeeds.csv")
breeding<-br%>%filter(lat>=-2.190181)
breeding<-breeding%>%group_by(id)%>%mutate(avg_speed=mean(Speed_km_h))
breed<-breeding%>%group_by(id)%>%filter(row_number()==1)
breed<-breed%>%mutate(avg_speed_m_s=avg_speed*5/18)
breed<-breed%>%dplyr::select(2,17:18)
names(breed) <- c("id","Breeding_avg_sp_km_hr", "Breeding_avg_sp_m_s")
speeds_section<-speeds_section%>%dplyr::select(1:4, 6:8)
speeds_section<-speeds_section%>%left_join(breed, by="id")
speeds_section<-speeds_section%>%dplyr::select("id","Bottom_avg_sp_km_hr","Middle_avg_sp_km_hr","Top_avg_sp_km_hr","Breeding_avg_sp_km_hr","Bottom_avg_sp_m_s","Middle_avg_sp_m_s","Top_avg_sp_m_s","Breeding_avg_sp_m_s")
write.csv(speeds_section,"/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/avgspeeds_all_Sections.csv")
#I have a graph of the speeds in the above worksheet
```
Graphing the sections
```{r}

all<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/allspR.csv")
all$location <- factor(all$location, levels = c("WAP - Cape Horn", "Cape Horn - Peninsula de Paracas", "Peninsula de Paracas- Salinas", "Salinas and above"))
ggplot(data=all)+
  geom_point(mapping=aes(x=location,y=speed), alpha=.25, color="blue",size=5)+ theme(axis.text.x=element_text(angle = -25, hjust = 0))+ylab("Speed (km/hr)")+xlab("Segment of Route")+ggtitle("Speed as a function of location")+ theme(plot.title = element_text(hjust=0.5))
#ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/speeds_Sections_all.jpg", width=8)


graph<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/speedsectionsforgraph.csv")
graph$Location <- factor(graph$Location, levels = c("WAP - Cape Horn", "Cape Horn - Peninsula de Paracas", "Peninsula de Paracas- Salinas", "Salinas and above"))
graph$ID<-as.factor(graph$ID)
ggplot(data=graph)+
  geom_line(mapping=aes(x=Location,y=Speed..km.hr., color=ID), size=1)

graph2<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/avgspeedgraphnumbers.csv")
graph2$ID<-as.factor(graph2$ID)
ggplot(data=graph2)+
  geom_line(mapping=aes(x=Location,y=Speed..km.hr., color=ID), size=1)+ylab( "Speed (km/hr)")
#ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/speeds_all_Sections_5.jpg", width=8)
```
Percentage of Routes in National waters
1) Get just those 5 whales, the bsam version, and split them into 5 separate data sheets
2) for each, check how many points DURING migration they have
3) For each, check how many points within each country
```{r}
all<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/bsam models/bsamspeeds_migonly.csv")
#we are going to test that it is migrators only
pp<-c(mean(all$lon),-40)
m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_point(data=all, aes(x=lon, y=lat),size=.5) + scale_color_discrete(guide="none") + mytheme
```
looks good, now we will get those migrators who went the whole way only
(Ptt= 112699, 121210, 123232, 131130, 166123)
```{r}
a112699<-all%>%filter(id==112699)
a121210<-all%>%filter(id==121210)
a123232<-all%>%filter(id==123232)
a131130<-all%>%filter(id==131130)
a166123<-all%>%filter(id==166123)

#graph individually
ggmap(m)+geom_point(data=a112699, aes(x=lon, y=lat),size=.5) + scale_color_discrete(guide="none") + mytheme
ggmap(m)+geom_point(data=a121210, aes(x=lon, y=lat),size=.5) + scale_color_discrete(guide="none") + mytheme
ggmap(m)+geom_point(data=a123232, aes(x=lon, y=lat),size=.5) + scale_color_discrete(guide="none") + mytheme
ggmap(m)+geom_point(data=a131130, aes(x=lon, y=lat),size=.5) + scale_color_discrete(guide="none") + mytheme
ggmap(m)+geom_point(data=a166123, aes(x=lon, y=lat),size=.5) + scale_color_discrete(guide="none") + mytheme

```

112699=98
  chile-47 pts = 48%
  peru-27 pts = 28%
  ecuador-3 = 3%
  =77/98 = 79%
  
121210=110
  chile = 43 = 39%
  peru =23 pts = 21%
  ecuador =4 = 4%
  total=64%
  
123232=100
  chile = 39 = 39%
  peru- 25 pts = 25%
  ecuador- 6= 6%
  total=70%
  
131130=104
  chile -55=53%
  peru- 18 = 17%
  ecuador - 3 = 3%
  total= 73%

166123=84
 chile - 33 - 39%
  peru- 25 - 30%
  ecuador-6 - 7%
  total-63- 75%
  
  72% total time

```{r}

p0<-ggmap(m)+geom_point(data=a112699, aes(x=lon, y=lat, col="red"),size=.05) + scale_color_discrete(guide="none") + mytheme
p1<-ggmap(m)+geom_point(data=a121210, aes(x=lon, y=lat, col="red"),size=.05) + scale_color_discrete(guide="none") + mytheme
p2<-ggmap(m)+geom_point(data=a123232, aes(x=lon, y=lat, col="red"),size=.05) + scale_color_discrete(guide="none") + mytheme
p3<-ggmap(m)+geom_point(data=a131130, aes(x=lon, y=lat, col="red"),size=.05) + scale_color_discrete(guide="none") + mytheme
p4<-ggmap(m)+geom_point(data=a166123, aes(x=lon, y=lat, col="red"),size=.05) + scale_color_discrete(guide="none") + mytheme


library("rgdal") # for `ogrInfo()` and `readOGR()`
library("tools") # for `file_path_sans_ext()`
library(ggplot2)
library("sp") # for `point.in.polygon()` and `spDists()`
library("tidyr") # for `gather()`
library("readr") # for `write_tsv()`
#path.eez.world.v8 <- ("/Users/danielpalacios/Documents/DMP/dmp_data/World_EEZ/World_EEZ_v8_20140228_LR")
path.eez.world.v9 <- ("/Users/michellemodest/Google Drive/Documents/UCSC/WWF/World_EEZ_v9_20161021_LR")
#fnam.eez.world.v8 <- "World_EEZ_v8_2014.shp"
fnam.eez.world.v9 <- "eez_lr.shp"
#eez.world.v8 <- readOGR(dsn = path.eez.world.v8, 
#                        layer = file_path_sans_ext(fnam.eez.world.v8))
eez.world.v9 <- readOGR(dsn = path.eez.world.v9, 
                        layer = file_path_sans_ext(fnam.eez.world.v9))
# A Large SpatialLinesDataFrame object with 281 features and 23 fields (18.9 Mb)

# Extract the EEZ for the USA:
#dat.eez.usa2 <- eez.world.v8[eez.world.v8@data$Country == "United States", ]
# For v. 9 use $Territory1 instead of $Country:
dat.eez.chile <- eez.world.v9[eez.world.v9@data$Territory1 == "Chile", ]
dat.eez.peru <- eez.world.v9[eez.world.v9@data$Territory1 == "Peru", ]
dat.eez.ec <- eez.world.v9[eez.world.v9@data$Territory1 == "Ecuador", ]
dat.eez.col <- eez.world.v9[eez.world.v9@data$Territory1 == "Colombia", ]
dat.eez.is <- eez.world.v9[eez.world.v9@data$Territory1 == "Islas San Félix and San Ambrosio", ]
dat.eez.an <- eez.world.v9[eez.world.v9@data$Territory1 == "Antarctica", ]
# A Formal class Large SpatialPolygonsDataFrame

# Fortify the shapefile data:
#dat.eez.usa2 <- fortify(dat.eez.usa2)
# `fortify.shape()` did not work for v. 8 so we had to use `fortify()`.
# message: Regions defined for each Polygons
dat.eez.chile <- fortify(dat.eez.chile)
dat.eez.peru <- fortify(dat.eez.peru)
dat.eez.ec <- fortify(dat.eez.ec) 
dat.eez.col <- fortify(dat.eez.col) 
dat.eez.is<- fortify(dat.eez.is)
dat.eez.an<- fortify(dat.eez.an) 
# a 10298x30 dataframe
 p0+ geom_path(data = dat.eez.chile, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.peru, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.col, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
      geom_path(data = dat.eez.ec, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.is, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
    geom_path(data = dat.eez.an, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)
 
 #ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/112699borders.svg", width=10)
  p1+ geom_path(data = dat.eez.chile, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.peru, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.col, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
      geom_path(data = dat.eez.ec, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.is, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
    geom_path(data = dat.eez.an, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)
  
  #ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/121210borders.svg", width=10)

    p2+ geom_path(data = dat.eez.chile, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.peru, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.col, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
      geom_path(data = dat.eez.ec, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.is, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
    geom_path(data = dat.eez.an, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)
  
  #ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/123232borders.svg", width=10)

    

    p3+ geom_path(data = dat.eez.chile, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.peru, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.col, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
      geom_path(data = dat.eez.ec, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.is, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
    geom_path(data = dat.eez.an, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)
  
 #ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/131130borders.svg", width=10)
    p4+ geom_path(data = dat.eez.chile, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.peru, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.col, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
      geom_path(data = dat.eez.ec, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.is, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
    geom_path(data = dat.eez.an, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)
    
  ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/Sample statistics/166123borders.svg", width=10)

ggmap(m)+geom_point(data=mig7, aes(x=x, y=y,col=Behavior,group=paste(Animal,Track)),alpha=0.7, size = 0.01)   + mytheme+ geom_path(data = dat.eez.chile, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.peru, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.col, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
      geom_path(data = dat.eez.ec, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.is, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
    geom_path(data = dat.eez.an, 
            aes(x = long, y = lat, group = group), 
            colour = "red", size = 0.3)
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/beh_map_points_borders.svg",height=20,width=25)
```