---
title: "Untitled"
output: html_document
---
#First attempt - Ben's model but with some updates (i.e. the t distribution, etc)
## first step is to wrangle up the data correctly
we want
1. only migrators
2. No duplicates (timestamp duplicates or location duplicates)
3. Get rid of NAs
4. Get rid of Z class
5. Get rid of pre-migration data
6. Calculate step lengths and tracks

##load the proper programs
```{r}
library(tidyverse)
library(lubridate)
library(knitr)
opts_chunk$set(echo=F,warning=F,message=F,fig.width = 11,fig.height = 5,cache=F)
library(boot)
library(raster)
library(tidyr)
library(ggplot2)
#library(MASS) This fucks with dplyr, load later
library(ggmap)
library(dplyr)
library(chron)
library(gridExtra)
library(stringr)
library(R2jags)
library(maptools)
library(reshape2)
#New model to be run, flag turned off if just updating.
newModel<-T

##ggplot theme
mytheme<-theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),axis.title.x=element_blank(),axis.title.y=element_blank(),panel.grid=element_blank())
```

## Only migrators
We have to decide whether we are keeping 123236 and 121208. Right now they are still included
```{r}
migrators<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/data compilations/all_animals/all_sex_nodup.csv")
migrators$Ptt<-as.factor(migrators$Ptt)
summary(migrators)
migrators<-migrators%>%filter(Ptt%in%c("112699", "121207", "121208", "121210", "123224", "123232", "123236", "131127", "131130", "131132", "131133", "131136"))
#test that it worked
mig<-migrators%>%
  group_by(Ptt)%>%
  filter(row_number()==1)
mig
migrators$timestamp<-as.POSIXct(migrators$timestamp,format="%Y-%b-%d %H:%M:%S",tz="GMT")

```
##2. No duplicates (timestamp duplicates or location duplicates)
```{r}
#first we will check if there are duplicate timestamps for any individuals
time<-migrators %>% group_by(Ptt,timestamp) %>% summarize(n=n()) %>% filter(n>1)
#looks like yes 
# Now lets test if there are any duplicate locations for any of the animals
location<-migrators%>%group_by(Ptt, Latitude.x, Longitude.x)%>%summarize(n=n())%>%filter(n>1)
#no duplicates!

ungroup(migrators)

migrators<-migrators[!is.na(migrators$timestamp),]

```

##4. Get rid of Z class

```{r}
migrators<-migrators%>%filter(Quality!="Z")
mig<-droplevels(migrators)
```

combine the genetic sex and pregnancy
```{r}
mig$Genetic.Sex<-as.character(mig$Genetic.Sex)
mig$Pregnant<-as.character(mig$Pregnant)
mig<-mig %>% 
  unite(Sex_Rep, Genetic.Sex, Pregnant)
test2<-mig%>%group_by(Sex_Rep)%>%summarize(n=n())%>%filter(n>1)
```

```{r}
mig2<-mig
mig2$Day<-days(mig2$timestamp)
mig2$Year<-years(mig2$timestamp)

mig2<-mig2%>%
  mutate(Year=mig2$Year)%>%
  mutate(Month=mig2$Month)%>%
  mutate(Day=mig2$Day,Date=format(timestamp,"%Y-%m-%d"))
mig2<-mig2%>%select(-Year, -Day)
mig2$Date<-ymd(mig2$Date)

f_l<-mig2%>%group_by(Ptt)%>%filter(row_number()==1 | row_number()==n())

na112699<-mig2%>%filter(Ptt=="112699")
fn121207<-mig2%>%filter(Ptt=="121207")
fn121208<-mig2%>%filter(Ptt=="121208")
m121210<-mig2%>%filter(Ptt=="121210")
fp123224<-mig2%>%filter(Ptt=="123224")
na123232<-mig2%>%filter(Ptt=="123232")
fn123236<-mig2%>%filter(Ptt=="123236")
na131127<-mig2%>%filter(Ptt=="131127")
fn131130<-mig2%>%filter(Ptt=="131130")
m131132<-mig2%>%filter(Ptt=="131132")
m131133<-mig2%>%filter(Ptt=="131133")
na131136<-mig2%>%filter(Ptt=="131136")

na112699<-na112699%>%filter(Date>=as.Date("2012-06-15"))
fn121207<-fn121207%>%filter(Date>=as.Date("2013-05-07"))
fn121208<-fn121208%>%filter(Date>=as.Date("2013-02-12"))

m121210<-m121210%>%filter(Date>=as.Date("2013-04-30"))
fp123224<-fp123224%>%filter(Date>=as.Date("2013-05-23"))

na123232<-na123232%>%filter(Date>=as.Date("2013-04-25"))

fn123236<-fn123236%>%filter(Date>=as.Date("2013-03-16"))
na131127<-na131127%>%filter(Date>=as.Date("2016-07-15"))
fn131130<-fn131130%>%filter(Date>=as.Date("2016-04-27"))
m131132<-m131132%>%filter(Date>=as.Date("2016-05-09"))
m131133<-m131133%>%filter(Date>=as.Date("2016-07-05"))
na131136<-na131136%>%filter(Date>=as.Date("2016-06-30"))

mig_migonly <- rbind(na112699,fn121207,fn121208,m121210,fp123224,na123232,fn123236,na131127,fn131130, m131132,m131133,na131136)


#Just an aside here

f_l_migration<-mig_migonly%>%group_by(Ptt)%>%filter(row_number()==1 | row_number()==n())


```

