---
title: "Correct R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##load the proper programs
```{r}
library(tidyverse)
library(lubridate)
library(knitr)
opts_chunk$set(echo=F,warning=F,message=F,fig.width = 11,fig.height = 5,cache=F)
library(boot)
library(raster)
library(tidyr)
library(ggplot2)
# MASS This fucks with dplyr, load later
library(ggmap)
library(dplyr)
library(chron)
library(gridExtra)
library(stringr)
library(R2jags)
library(maptools)
library(reshape2)
#New model to be run, flag turned off if just updating.
newModel<-T
register_google(key="<insertgooglekeyhere>")
```

## Only migrators
We have to decide whether we are keeping 123236 and 121208. Right now they are still included
NEW MIGRATORS!!
```{r}
migrators<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/data compilations/All_animals_cleaned.csv")
migrators$Ptt<-as.factor(migrators$Ptt)
summary(migrators)
#This DOES have the 2016 tags that were on and off, 131133", "131136". Probably want to remove later 
migrators<-migrators%>%filter(Ptt%in%c("112699", "121207", "121210", "123224", "123232", "123236", "131127",  "131130", "131132", "131133","131136","166122", "166123", "166125", "166126", "166128"))

#test that it worked
mig<-migrators%>%
  group_by(Ptt)%>%
  filter(row_number()==1)
mig
```
```{r}
migrators$timestamp<-as.character(migrators$timestamp)
migrators$timestamp<-as.POSIXct(migrators$timestamp,format="%Y-%m-%d %H:%M:%S",tz="GMT")
```

```{r}
#first we will check if there are duplicate timestamps for any individuals
time<-migrators %>% group_by(Ptt,timestamp) %>% summarize(n=n()) %>% filter(n>1)
#looks like no. 
# Now lets test if there are any duplicate locations for any of the animals
location<-migrators%>%group_by(Ptt, Latitude, Longitude)%>%summarize(n=n())%>%filter(n>1)
#but thats ok we keep it
ungroup(migrators)
```

##3. Get rid of NAs
```{r}
#remove empty timestamps
migrators<-migrators[!is.na(migrators$timestamp),]#there aren't any
```



##4. Get rid of Z class

```{r}
migrators<-migrators%>%filter(Quality!="Z")
mig<-droplevels(migrators)
```



##5. Get rid of pre-migration data
```{r}
mig2<-mig
#test
mig_migonly<-mig2


mig2$Date<-ymd(mig2$Date)

f_l<-mig2%>%group_by(Ptt)%>%filter(row_number()==1 | row_number()==n())

#Just an aside here

f_l_migration<-mig_migonly%>%group_by(Ptt)%>%filter(row_number()==1 | row_number()==n())


```
Let's graph things
```{r}
pp<-c(mean(mig2$Longitude),-40)
mig2$Year<-as.factor(mig2$Year)

m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_point(data=mig2, aes(x=Longitude, y=Latitude, col=Year),size=.25)  + facet_wrap(~Ptt,ncol=3) 
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/individualspremodel2.jpg", width=8, height=30,  units="in")


biopsy<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/data compilations/Palmer data/Mn_sex.csv")

biopsy_prim<-biopsy%>%
  count(Tag.ID)%>%
  filter(n>1)

mig2$Ptt<-as.factor(mig2$Ptt)
Whale_sex<-mig2%>%
  left_join(biopsy, c("Ptt"="Tag.ID"))%>%dplyr::select(Ptt:Longitude.x, timestamp:Year.x,Genetic.Sex:Pregnant)

names(Whale_sex)[names(Whale_sex)=='Year.x']<-'Year'
names(Whale_sex)[names(Whale_sex)=='Latitude.x']<-'Latitude'
names(Whale_sex)[names(Whale_sex)=='Longitude.x']<-'Longitude'
Mig_sex<-Whale_sex

Mig_sex<-Mig_sex%>%
  unite(Sex_Pregnancy,Genetic.Sex,Pregnant, remove=FALSE)
Mig_sex$Ptt<-as.factor(Mig_sex$Ptt)
Mig_sex2<-Mig_sex%>%arrange(desc(Sex_Pregnancy),.by_group=TRUE)
#now graph again
pp<-c(mean(mig2$Longitude),-40)
Mig_sex$Sex_Pregnancy<-as.factor(Mig_sex$Sex_Pregnancy)
m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_point(data=Mig_sex2, aes(x=Longitude, y=Latitude, col=Sex_Pregnancy),size=.25)  + facet_wrap(~Ptt,nrow=3) 
#ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/individualspremodel.jpg", width=8, units="in")

#Now we will plot all at once
pp<-c(mean(mig2$Longitude),-40)
Mig_sex2$Sex_Pregnancy<-as.factor(Mig_sex2$Sex_Pregnancy)
m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_point(data=Mig_sex, aes(x=Longitude, y=Latitude, col=Sex_Pregnancy),size=.15)
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/allpremodel_sex.jpg", width=8, units="in")
```

##6. Calculate step lengths and tracks
First we are going to match the names and labels from Ben's code
```{r}
#standardize column names to match the simulation
#Create an animal tag.
str(mig_migonly)
mig_migonly<-mig_migonly%>%dplyr::select(1,3,5,7:9,21:24)
names(mig_migonly) <- c("X", "Animal", "Date", "argos.lc", "y", "x", "timestamp", "Month", "Day", "Year" )
mig3<-mig_migonly
#rearrange
mig3<-mig3%>%dplyr::select("X", "Animal", "argos.lc", "x", "y", "timestamp", "Month", "Year", "Date")

```

Ok now we are really going off the deep-end with Ben's code, so starting a new name so that mig4 is the thing we can always go back to
```{r}
#holding mig 4 for later
mig4<-mig3


##ggplot theme
mytheme<-theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(),axis.text.y=element_blank(),axis.ticks.y=element_blank(),axis.title.x=element_blank(),axis.title.y=element_blank(),panel.grid=element_blank())


#remake spatial object
d<-SpatialPointsDataFrame(cbind(mig3$x,mig3$y),data=mig3,proj4string=CRS("+proj=longlat +datum=WGS84"))

#Plot
pp<-c(mean(mig3$x),-40)
m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_point(data=mig3, aes(x=x, y=y,col=as.factor(Animal)),size=.25) + scale_color_discrete(guide="none") + mytheme

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/color_map.jpg")

mig3$Month<-factor(mig3$Month,levels=month.name)
m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mig3, aes(x=x, y=y,col=as.factor(Animal)),size=.5)  + facet_wrap(~Month,nrow=2) + scale_color_discrete(guide="none") + theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank()) + labs(x="",y="")
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/color_map_month.jpg")
ggmap(m)+geom_point(data=mig3, aes(x=x, y=y,col=as.factor(Animal)),size=.5)  + facet_wrap(~Animal, nrow=2) + scale_color_discrete(guide="none") + theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank()) + labs(x="",y="")

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/color_map_ind.jpg")
#black and white
ggplot()+geom_path(data=mig3, aes(x=x, y=y,group=Animal),size=.5)  + facet_wrap(~Month,nrow = 2)  + borders(fill="grey90") + coord_cartesian(ylim = c(-70,10),xlim=c(-60,-85)) + theme_bw() + mytheme + scale_fill_manual(values=rep("NA",8)) + scale_color_discrete(guide="none")
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/bw_map_month.jpg")
```

```{r}
##Time is the beginning of the first point.
step_length=12

dumb1<-split(mig3,mig3$Animal)

#time diff function
timed<-function(d,step_length){
  d$j[1]<-0
  for (x in 2:nrow(d)){
    d$j[x]<-as.numeric(difftime(as.POSIXct(d$timestamp[x]),as.POSIXct(d$timestamp[x-1]),units="mins"))/(step_length*60)
  }
  
  #Split out track endings
  ends<-c(1,which(d$j>1),nrow(d))

  for(w in 2:length(ends)){
    d[ends[w-1]:ends[w],"Track"]<-w-1
  }
  
  #remove tracks that are shorter than 24 hrs
  track_time<-d %>% group_by(Track) %>% summarize(mt=difftime(max(as.POSIXct(timestamp)),min(as.POSIXct(timestamp)),units="hours")) %>% filter(mt>=24) %>% .$Track
  
  d<-d[d$Track %in% track_time,]
  
  #renumber the tracks
  d$Track<-as.numeric(as.factor(d$Track))
  return(d)
  }

dumb2<-lapply(dumb1,timed,step_length=step_length)

#Format matrices for jags
mig3<-bind_rows(dumb2)

######recode whales
#mig3$Animal<-as.numeric(as.factor(mig3$Animal))

dumb3<-split(mig3,list(mig3$Animal,mig3$Track),drop=TRUE)

dumb4<-lapply(dumb3,function(x){
#How many observations in each step length segment
x$step<-as.numeric(cut(as.POSIXct(x$timestamp),paste(step_length,"hours")))
return(x)
})

mig3<-bind_rows(dumb4)

#refactor animal
mig3$Ptt<-mig3$Animal
mig3$Animal<-as.numeric(as.factor(mig3$Animal))


```
Now we go and do it my way
```{r}
mig3<-mig3 %>% group_by(Animal,Track,step) %>% mutate(jStep=1:n())

mig4<-mig3%>%mutate(j=ifelse(step == 1 & jStep == 1,0,j))  
splitmig<-split(mig4,list(mig4$Animal,mig4$Track),drop=TRUE)
                    
                    splitmigsum<-lapply(splitmig,function(x){
                      x$sumj<-cumsum(x$j)
                      return(x)
                    })
                    
                    mig5<-bind_rows(splitmigsum)

                    revtrunc <- function(x) { sign(x) * (x - floor(x)) } 
mig5$basej<-revtrunc(mig5$sumj)                    

mig5<-mig5%>%mutate(step_new=floor(sumj)+1)

#gonna have to redo jstep too
mig5<-mig5 %>% group_by(Animal,Track,step_new) %>% mutate(jStep_new=1:n())
ungroup(mig5)
mig6<-mig5%>%dplyr::select(Ptt, Animal, argos.lc, x, y, timestamp, Month, Year, Date, Track, step_new, jStep_new, basej)

names(mig6)[names(mig6) == 'step_new'] <- 'step'
names(mig6)[names(mig6)== 'jStep_new']<-'jStep'
names(mig6)[names(mig6)== 'basej']<-'j'
head(mig6)
```
```{r}


mig6$Month<-factor(mig6$Month,levels=month.name)
m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_point(data=mig6, aes(x=x, y=y,col=as.factor(Animal)),size=.2)  + facet_wrap(~Ptt,nrow=2) + scale_color_discrete(guide="none") + theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank()) + labs(x="",y="")
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/color_map_ind_afterfilteredandremoved.jpg")

ggmap(m)+geom_path(data=mig6, aes(x=x, y=y,col=as.factor(Animal)),size=.5)  + facet_wrap(~Month,nrow=2) + scale_color_discrete(guide="none") + theme(axis.text.x=element_blank(),axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank()) + labs(x="",y="")
```

```{r}
#total number of steps per track/animal
steps_all<-mig6 %>% group_by(Animal,Track) %>% summarize(n=length(unique(step)))


#Cast time array
j<-acast(mig6,Animal~Track~step~jStep,value.var="j")

#how many observations per individual in each step. Basically the j step. 
mig6$step<-factor(mig6$step,levels=1:max(steps_all$n))
idx<-melt(table(mig6$Animal,mig6$Track,mig6$step))
colnames(idx)<-c("Animal","Track","step","jStep")
idx<-acast(data=idx,Animal~Track~step) # what this is saying is that
#we get 225 matrices, 1 for each step, rownames are animal number, column names are track #, values are the number of observations in that animal and track at that step

#month array
mig6$MonthF<-as.numeric(factor(mig6$Month,levels=month.name))

MonthA<-acast(mig6,Animal~Track~step,value.var="MonthF",fun.aggregate = min)
MonthA[!is.finite(MonthA)]<-NA

#Individuals
ind=length(unique(mig6$Animal))

#tracks per indivudal
tracks<-mig6 %>% group_by(Animal) %>% summarize(tracks=length(unique(Track))) %>% .$tracks

#steps per track
#resulting matrix - rows are Animal ID, column names are Track number, each spot is number of steps
steps<-acast(steps_all,Animal~Track,value.var="n")

#index array for merging
mig6$Index<-1:nrow(mig6)

#obs array
obs<-melt(mig6,measure.vars=c("x","y"))
obs<-acast(obs,Animal~Track~step~jStep~variable)

MonthIndex<-acast(mig6,Animal~Track~step,value.var="Month",fun=function(x) names(which.max(table(x))) )

#argos error class array
mig6$argos.lc<-factor(mig6$argos.lc,levels=c(3,2,1,0,"A","B"))
mig6$numargos<-as.numeric(mig6$argos.lc)
obs_class<-acast(mig6,Animal~Track~step~jStep,value.var="numargos")
```

```{r}
#Plot
head(mig6)
mig6<-mig6 %>% arrange(Animal,Track,step,jStep)

pp<-c(mean(mig6$x),mean(mig6$y))
m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=mig6, aes(x=x, y=y,col=as.factor(Animal),group=paste(Animal,Track)),size=.5) + scale_color_discrete(guide="none") + mytheme

ggplot() +  geom_path(data=mig4,col='blue',aes(x=x,y=y)) +  geom_path(data=mig6,col='red',aes(x=x,y=y),size=2,alpha=0.7)  + facet_wrap(~Animal,scales="free",ncol=4)

ggplot(data=steps_all,aes(x=n*step_length/24)) + geom_histogram() + labs(x="Days") + ggtitle("Track Length")
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/track_lengths.jpg")
ggplot(data=steps_all,aes(x=Track)) + geom_histogram() + labs(x="Subtracks per Animal")
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/tracks_animal.jpg")
write.csv(mig6, "")
```

```{r}
library(MASS)
#source jags file
source("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/JAGS_code/Bens_model.R")

#prior cov shape
#the left is the jags and the right is the R
R <- diag(c(1,1))
data=list(argos=obs,steps=steps,R=R,ind=ind,j=j,idx=idx,tracks=tracks,argos_class=obs_class,Month=MonthA,Months=max(MonthA,na.rm=T))

#parameters to track
pt<-c("theta","gamma","phi","alpha","state","gamma_mu","gamma_tau")

if(newModel){
  system.time(jagM<-jags.parallel(model.file = "/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/JAGS_code/Bens_model.R",data=data,n.chains=2,parameters.to.save=pt,n.iter=270000,n.burnin=250000,n.thin=8,DIC=FALSE))
}
```



Chains
```{r}

#extract parameter name
#what this does is it creates a news column called parameter that just matches the parameter name from par column
pc12<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/pc12no16.csv")
pc12<-pc12%>%dplyr::select(2:5)
pc12$parameter<-data.frame(str_match(pc12$par,"(\\w+)"))[,-1]

#Extract index
#Now we are splitting pc12 along parameter lines into groups
splitpc12<-split(pc12,pc12$parameter)

#single index
#I think this means we are taking the theta list only and we are adding the behavior
splitpc12[c("theta")]<-lapply(splitpc12[c("theta")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+)]"))[,3]
    pc12<-data.frame(x,Behavior=sv)
    return(pc12)
})

##Double index
#single index
splitpc12[c("alpha","gamma")]<-lapply(splitpc12[c("alpha","gamma")],function(x){
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+)]"))[,3:4]
    colnames(sv)<-c("Behavior","Month")
    pc12<-data.frame(x,sv)
    return(pc12)
})

#Three index
#Adding Animal, Track, step and behavior to Phi
splitpc12[c("phi")]<-lapply(splitpc12[c("phi")],function(x){
#As matrices
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+),(\\d+)]"))[,3:6]
    colnames(sv)<-c("Animal","Track","step","Behavior")
    pc12<-data.frame(x,sv)
})


#State index
#adding Animal, Track and step to state
splitpc12[c("state")]<-lapply(splitpc12[c("state")],function(x){
#As matrices
    sv<-data.frame(str_match(x$par,"(\\w+)\\[(\\d+),(\\d+),(\\d+)]"))[,3:5]
    colnames(sv)<-c("Animal","Track","step")
    pc12<-data.frame(x,sv)
})

#bind all matrices back together
pc12<-bind_rows(splitpc12) 
# some warning about unequal factor levels
rm(splitpc12)

#associate month with state
#first you have to bring back the month index
mm<-melt(MonthIndex)
colnames(mm)<-c("Animal","Track","step","MonthA")
pc12<-merge(pc12,mm,by=c("Animal","Track","step"),all.x=T)

#complete level matching, ugly but successful

pc12[pc12$parameter %in% c("alpha","gamma"),]$MonthA<-month.name[as.numeric(pc12$Month[pc12$parameter %in% c("alpha","gamma")])]
pc12$MonthA<-factor(pc12$MonthA,levels=month.name)
```
Just a summary at this point
alpha-> behavior, month
gamma-> behavior, month
theta->behavior
phi-> Animal, track, step, behavior
state-> animal, track, step

lso when you bind the rows together there is a warning about unequal factor levels
HAVING ISSUES HERE, won't let me run the last bit - says 13 variables to replace 12 variablves

NEvermind....now it is working - maybe I ran it twice last time or something?

Plot all but phi and state
I THINK THIS IS THE CHAINS BUT WE NEED A BETTER WAY OF VISUALIZING ALSO I FEEL LIKE WE NEED MORE DRAWS
```{r}
#plot all but phi and state
#I think this is the chains but we need a better way of visualizing
ggplot(pc12[pc12$parameter %in% c("alpha"),],aes(x=Draw,y=value)) + facet_wrap(~par,ncol=3) + geom_line(aes(col=as.factor(chain))) + labs(col="Chain")

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/alpha_chains.jpg", height=10, width=10)

ggplot(pc12[pc12$parameter %in% c("gamma"),],aes(x=Draw,y=value)) + facet_wrap(~par,ncol=3) + geom_line(aes(col=as.factor(chain))) + labs(col="Chain")

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/gamma_chains.jpg", height=10, width=10)

ggplot(pc12[pc12$parameter %in% c("theta"),],aes(x=Draw,y=value)) + facet_wrap(~par,ncol=3) + geom_line(aes(col=as.factor(chain))) + labs(col="Chain")

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/theta_chains.jpg")




a<-pc12 %>% filter(parameter %in% c("state","phi")) %>% filter((parameter == "phi"&Behavior==2)|parameter =="state") %>% dplyr::select(Draw,chain,Animal,Track,step,value,parameter) %>% spread(key=parameter,value=value)
ggplot(a,aes(x=phi,fill=as.factor(state-1))) + geom_density(alpha=0.7) + labs(fill="Behavior")
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/behavior.jpg")
```

Write some important csv files
```{r}
#write.csv
write.csv(pc12,"/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/outdata_parse.csv")
pc12<-read.csv("/Users/michellemodest/Dropbox/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/outdata_parse.csv")
pc12$Behavior<-as.factor(pc12$Behavior)
```
Plot
```{r}
#Plot
ggplot(data=pc12[pc12$parameter %in% c("gamma"),],aes(x=value)) + geom_histogram() + facet_wrap(~par,ncol=3) + theme_bw()  + ggtitle("Estimated parameters")

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/gamma_hist.jpg", width=10, height=10)

ggplot(data=pc12[pc12$parameter %in% c("alpha"),],aes(x=value)) + geom_histogram() + facet_wrap(~par,ncol=3) + theme_bw()  + ggtitle("Estimated parameters")

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/alpha_hist.jpg", height=10, width=10)

ggplot(data=pc12[pc12$parameter %in% c("theta"),],aes(x=value)) + geom_histogram() + facet_wrap(~par,ncol=3) + theme_bw()  + ggtitle("Estimated parameters")

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/theta_hist.jpg")


```
Change in Autocorrelation over time
```{r}
pc12$MonthA<-factor(pc12$MonthA,levels=month.name)
g<-pc12[pc12$parameter %in% c("gamma"),] %>% mutate(bstate=ifelse(Behavior==1,"Traveling","Area-restricted Search"))
ggplot(data=g,aes(x=value,fill=MonthA)) + geom_density(alpha=0.7) + theme_bw()+ facet_wrap(~bstate,scales="free") + labs(fill="Month",x="Degree of movement autocorrelation") + scale_fill_brewer(palette="RdBu")

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/Gammaplot.jpg",height=3,width=6,dpi=300)

a<-pc12[pc12$parameter %in% c("gamma"),] %>% mutate(bstate=ifelse(Behavior==1,"Traveling ","Foraging"))

ggplot(data=a,aes(y=value,x=MonthA)) + geom_boxplot() + theme_bw()  + ggtitle("Movement autocorrelation") + labs(x="Month",fill="Behavior") + facet_wrap(~bstate,scales="free")+ theme(axis.text.x=element_text(angle = -90, hjust = 0))
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/Gammabox.jpg",height=3,width=8)
```
WE seem to BE MISSING SO MUCH INFO IN G- like months, animals, etc
Also two month A's are part of the problem - maybe that was the 13th parameter it was talking about
Also the months have to be factors
Change in transition probabilities over time
```{r}
a<-pc12[pc12$parameter %in% c("alpha"),] %>% mutate(bstate=ifelse(Behavior==1,"Remain Traveling ","Switch to Traveling"))
ggplot(data=a,aes(x=value,fill=MonthA)) + geom_density(alpha=0.7) + theme_bw()+ facet_wrap(~bstate,scales="free") + labs(fill="Month",x="Transition Probability") + scale_fill_brewer(palette="RdBu")
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/Alphaplot.jpg",height=4,width=6,dpi=300)

ggplot(data=a,aes(y=value,x=MonthA)) + geom_boxplot() + theme_bw()+ facet_wrap(~bstate,scales="free") + labs(x="Month",y="Transition Probability")+ theme(axis.text.x=element_text(angle = -90, hjust = 0))
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/Alphabox.jpg",height=3,width=8)
```

Parameter Summary
(I took out all the monthA's, they need to go back)
```{r}
psum<-pc12 %>% filter(!parameter %in% c("phi","state")) %>% group_by(parameter,par) %>% summarize(mean=mean(value),lower=quantile(value,0.05),upper=quantile(value,0.95))

as.data.frame(psum)

write.csv(psum, "/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/paramsum.csv")

ggplot(psum,aes(x=par,y=mean,ymax=upper,ymin=lower)) + geom_linerange() + geom_point() + facet_wrap(~parameter,scales="free", ncol=2)+ theme(axis.text.x=element_text(angle = -90, hjust = 0))

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/parameter_mean.jpg", height=10, width=6)
#format gamma
gammadf<-pc12 %>% filter(parameter=="gamma") %>% group_by(MonthA,Behavior) %>% summarize(mean=mean(value))

write.csv(gammadf, "/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/gammasum.csv")
#format alpha
alphadf<-pc12 %>% filter(parameter=="alpha") %>% group_by(MonthA,Behavior) %>% summarize(mean=mean(value))
write.csv(alphadf, "/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/alphasum.csv")
#both
simdf<-pc12 %>% filter(parameter %in% c("gamma","alpha")) %>% group_by(MonthA,Behavior,parameter) %>% summarize(mean=mean(value))

```

Behavioral Prediction
```{r}
#mean transition probability
mean_phi<-pc12 %>% filter(parameter=="phi",Behavior==2) %>% mutate(step=as.numeric(as.character(step))) %>% group_by(Animal,Track,step) %>% summarize(phi=median(value)) %>% arrange(Animal,Track,step)
write.csv(mean_phi, "/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/mean_phi.csv")
#mode state for each position
mode_state<-pc12 %>% filter(parameter %in% "state") %>% mutate(bstate=ifelse(value==1,"Traveling","Area-restricted Search")) %>% group_by(Animal,Track,step,bstate) %>% summarize(n=n()) %>% spread(key=bstate,value=n,fill=0)%>% mutate(ratio=`Area-restricted Search`/Traveling) %>% mutate(state=as.character(ifelse(ratio>=1,"Area-restricted Search","Traveling")))

read.csv<-("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/mig6correct.csv")

mig6<-merge(mig6,mean_phi,by=c("step","Track","Animal"))
mig6<-merge(mig6,mode_state,by=c("step","Track","Animal"))

write.csv(mig6, "/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/mig6_withparmodel12.csv")
mig6<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/mig6_withparmodel12.csv")
#make factor
mig6$bstate<-factor(mig6$state)

#ensure order for plotting
mig6<-mig6 %>% arrange(Animal,Track,step,jStep)
#Now we are going to use the methods from the other papers and calculate the ratio
names(mig6)[names(mig6) == 'Area.restricted.Search'] <- 'ARS'
mig7<-mig6%>%mutate(Behavior_ratio=((ARS*2)+(Traveling*1))/5000)
mig7$Behavior<-cut(mig7$Behavior_ratio,c(.99,1.25,1.75,2.1),labels=c("Traveling","Unknown","ARS"))
  mig7<-mig7 %>% arrange(Animal,Track,step,jStep)
  write.csv(mig7, "/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/mig6_withunknowntoo7-4-2020.csv")
pp<-c(mean(mig7$x),-40)
m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="hybrid",color="bw")
ggmap(m)+geom_point(data=mig7, aes(x=x, y=y,col=Behavior,group=paste(Animal,Track)),size=.3,alpha=0.7)   + mytheme
#ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/animals_state_map_noborders.jpg", width=8)

pp<-c(mean(mig7$x),-40)
m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="hybrid",color="bw")
p0<-ggmap(m)+geom_path(data=mig7, aes(x=x, y=y,col=Behavior,group=paste(Animal,Track)),size=.5,alpha=0.7)   + mytheme


library("rgdal") # for `ogrInfo()` and `readOGR()`
library("tools") # for `file_path_sans_ext()`
library(ggplot2)
library("sp") # for `point.in.polygon()` and `spDists()`
library("tidyr") # for `gather()`
library("readr") # for `write_tsv()`
#path.eez.world.v8 <- ("/Users/danielpalacios/Documents/DMP/dmp_data/World_EEZ/World_EEZ_v8_20140228_LR")
path.eez.world.v9 <- ("/Users/michellemodest/Google Drive/Documents/UCSC/WWF/World_EEZ_v9_20161021_LR")
#fnam.eez.world.v8 <- "World_EEZ_v8_2014.shp"
fnam.eez.world.v9 <- "eez_lr.shp"
#eez.world.v8 <- readOGR(dsn = path.eez.world.v8, 
#                        layer = file_path_sans_ext(fnam.eez.world.v8))
eez.world.v9 <- readOGR(dsn = path.eez.world.v9, 
                        layer = file_path_sans_ext(fnam.eez.world.v9))
# A Large SpatialLinesDataFrame object with 281 features and 23 fields (18.9 Mb)

# Extract the EEZ for the USA:
#dat.eez.usa2 <- eez.world.v8[eez.world.v8@data$Country == "United States", ]
# For v. 9 use $Territory1 instead of $Country:
dat.eez.chile <- eez.world.v9[eez.world.v9@data$Territory1 == "Chile", ]
dat.eez.peru <- eez.world.v9[eez.world.v9@data$Territory1 == "Peru", ]
dat.eez.ec <- eez.world.v9[eez.world.v9@data$Territory1 == "Ecuador", ]
dat.eez.col <- eez.world.v9[eez.world.v9@data$Territory1 == "Colombia", ]
dat.eez.is <- eez.world.v9[eez.world.v9@data$Territory1 == "Islas San Félix and San Ambrosio", ]
dat.eez.an <- eez.world.v9[eez.world.v9@data$Territory1 == "Antarctica", ]
# A Formal class Large SpatialPolygonsDataFrame

# Fortify the shapefile data:
#dat.eez.usa2 <- fortify(dat.eez.usa2)
# `fortify.shape()` did not work for v. 8 so we had to use `fortify()`.
# message: Regions defined for each Polygons
dat.eez.chile <- fortify(dat.eez.chile)
dat.eez.peru <- fortify(dat.eez.peru)
dat.eez.ec <- fortify(dat.eez.ec) 
dat.eez.col <- fortify(dat.eez.col) 
dat.eez.is<- fortify(dat.eez.is)
dat.eez.an<- fortify(dat.eez.an) 
# a 10298x30 dataframe
 p0+ geom_path(data = dat.eez.chile, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.peru, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.col, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
      geom_path(data = dat.eez.ec, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.is, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
    geom_path(data = dat.eez.an, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)
 
  p0+ geom_path(data = dat.eez.chile, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.peru, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.col, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
      geom_path(data = dat.eez.ec, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.is, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
    geom_path(data = dat.eez.an, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)
ggmap(m)+geom_point(data=mig7, aes(x=x, y=y,col=Behavior,group=paste(Animal,Track)),alpha=0.7, size = 0.01)   + mytheme+ geom_path(data = dat.eez.chile, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.peru, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.col, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
      geom_path(data = dat.eez.ec, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
   geom_path(data = dat.eez.is, 
            aes(x = long, y = lat, group = group), 
            colour = "yellow", size = 0.3)+
    geom_path(data = dat.eez.an, 
            aes(x = long, y = lat, group = group), 
            colour = "red", size = 0.3)
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/beh_map_points_borders.svg",height=20,width=25)
```

#Spatial Prediction
```{r}
m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="hybrid",color="bw")

ggmap(m)+geom_path(data=mig6, aes(x=x, y=y,col=bstate,group=paste(Animal,Track)),size=.2) + mytheme + labs(col="Behavior") +  facet_wrap(~Month)

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/beh_month_map.svg",height=20,width=25)

ggmap(m)+geom_point(data=mig6, aes(x=x, y=y,col=bstate,group=paste(Animal,Track)),size=.2) + mytheme + labs(col="Behavior") +  facet_wrap(~Month)+ geom_path(data = dat.eez.chile, 
            aes(x = long, y = lat, group = group), 
            colour = "red", size = 0.3)+
   geom_path(data = dat.eez.peru, 
            aes(x = long, y = lat, group = group), 
            colour = "red", size = 0.3)+
   geom_path(data = dat.eez.col, 
            aes(x = long, y = lat, group = group), 
            colour = "red", size = 0.3)+
      geom_path(data = dat.eez.ec, 
            aes(x = long, y = lat, group = group), 
            colour = "red", size = 0.3)+
   geom_path(data = dat.eez.is, 
            aes(x = long, y = lat, group = group), 
            colour = "red", size = 0.3)+
    geom_path(data = dat.eez.an, 
            aes(x = long, y = lat, group = group), 
            colour = "red", size = 0.3)
mig7$Month<-factor(mig7$Month,levels=month.name)
ggmap(m)+geom_point(data=mig7, aes(x=x, y=y,col=Behavior,group=paste(Animal,Track)),size=.01)  + facet_wrap(~Month,nrow=2)+ theme(axis.text.x=element_text(angle = -90, hjust = 0))
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/animals_month_behavior.svg",height=20,width=25)



#Pub version
ggplot()+ theme_set(theme_gray(base_size = 18)) +geom_path(data=mig7, aes(x=x, y=y,col=Behavior,group=paste(Animal,Track)),size=0.7)  + borders(fill="grey40",size=0.5) + coord_cartesian(ylim = c(-69,10),xlim=c(-55,-85)) + theme_bw() + mytheme + scale_fill_manual(values=rep("NA",8)) + labs(col="Behavior") + scale_color_manual(values=c("deeppink","darkturquoise"))  + facet_wrap(~Month,nrow=2) + theme(legend.position="bottom") 

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/month_behavior_bw.jpg",height=7,width=10)
```

#By Individual

```{r}
ggmap(m)+geom_point(data=mig7, aes(x=x, y=y,col=Behavior,group=paste(Animal,Track)),size=.01)   + mytheme + facet_wrap(~Animal)

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/individuals_behavior_point.svg", height=20, width=25)

```

##Autocorrelation in Behavior
```{r}
#create proper time stamp
ggplot(data=mig7,aes(x=as.POSIXct(timestamp),y=as.factor(Behavior))) + geom_line(aes(group=Track),size=1)+ theme_bw()  + facet_wrap(~Ptt,scales="free") + labs(x="Time",y="Behavior") + theme(axis.text.x=element_text(angle = -90, hjust = 0))

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/individuals_timeline.jpg", width=25, height=10)

```
As single timeline
```{r}
ggplot(data=mig7,aes(x=as.POSIXct(timestamp),y=as.factor(Behavior))) + geom_line(aes(group=paste(Animal,Track)),size=1)+ theme_bw()  + labs(x="Time",y="Behavior") + facet_wrap(~Year,scales="free")
ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/behavior_year.jpg")
```

We would like to make the initial graphs for all the animals before track filtering, and including their sex
```{r}
biopsy<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/data compilations/Palmer data/Mn_sex.csv")

biopsy_prim<-biopsy%>%
  count(Tag.ID)%>%
  filter(n>1)


mig3$Animal<-as.factor(mig3$Animal)
Whale_sex<-mig3%>%
  left_join(biopsy, c("Animal"="Tag.ID"))%>%dplyr::select(Animal:Date, Genetic.Sex:Pregnant)
#lets see if we have sex info for more now

mig<-Whale_sex%>%
  group_by(Animal)%>%
  filter(row_number()==1)

names(Whale_sex)[names(Whale_sex)=='Year.x']<-'Year'


Mig_sex<-Whale_sex

Mig_sex<-Mig_sex%>%
  unite(sex.preg,Genetic.Sex,Pregnant, remove=FALSE)
flsex<-Mig_sex%>%
  group_by(Animal)%>%
  filter(row_number()==1)

m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_point(data=Mig_sex, aes(x=x, y=y,col=Genetic.Sex),size=.25)+labs(col="Genetic.Sex")

#Plot
pp<-c(mean(Mig_sex$x),-40)
m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_point(data=Mig_sex, aes(x=x, y=y,col=Genetic.Sex),size=.25)+ scale_color_manual(breaks = c("F", "M", "NA"),
                        values=c("pink", "blue", "yellow")) +labs(col="Genetic.Sex")

ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/JAGS Models/Model 12/4_13/correct version/MFmap.jpg")
```

