---
title: "Untitled"
output: html_document
---


```{r }
register_google(key="<insertgooglekeyhere>")
library(tidyverse)
library(onewaytests)
mdat<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/linear regression speed/id_jday_speed_sex.csv")

```
lets fix the sex fields
```{r}
mdat[10,4]<-"F"
mdat[10,5]<-"P"
mdat[12,5]<-"P"
mdat[13,5]<-"NP"
mdat[14,5]<-"P"
mdat2<-mdat%>%
  unite(sex.preg,Sex,Pregant, remove=FALSE)
mf<-lm(Mean_speed~sex.preg, mdat2)
summary(mf)
mdat3<-mdat2%>%filter(sex.preg!="_")
mf2<-lm(Mean_speed~sex.preg, mdat3)
summary(mf2)
```

First we are going to filter it out so that we are only using speeds in the actual migration

```{r}

```
Ok I think we are all good and that 6/02 might just be wrong in the other sheet. Worth looking at though.

```{r}
xdml<-lm(Mean_speed~Julian_Day, mdat)
summary(xdml)
```
what if I did M vs F
```{r}
mf<-lm(Mean_speed~Sex, mdat)
summary(mf)
```
NAAAH

Now we are going to look at the residuals
```{r}
plot(fitted(xdml),residuals(xdml))
qqnorm(residuals(xdml))
dfbeta(xdml)
hist(residuals(xdml))
#looks like animal 7, i.e. 131130 is influencing things rather a lot, lets try without
mdat2<-mdat%>%filter(ID!=131130)

norm2<-lm(Mean_speed~Julian_Day, mdat2)
summary(norm2)
plot(fitted(norm2),residuals(norm2))
qqnorm(residuals(norm2))
dfbeta(norm2)

```
Lets try log transofmring
```{r}


mdat$logspeed<-log10(mdat$Mean_speed)
xdml2<-lm(logspeed~Julian_Day, mdat)
summary(xdml2)
plot(fitted(xdml2),residuals(xdml2))
hist(residuals(xdml2))
scatter.smooth(x=mdat$Julian_Day, y=mdat$logspeed, main="logspeed ~ Julian_Day")  # scatterplot
dfbeta(xdml2)

#get rid of the outlier
mdat2$logspeed<-log10(mdat2$Mean_speed)
log2<-lm(logspeed~Julian_Day, mdat2)
summary(log2)
plot(fitted(log2),residuals(log2))
qqnorm(residuals(log2))
hist(residuals(log2))
dfbeta(log2)
```

Both of the models run better without 131130. So probably say that. 


BASICALLY
looks like we do the regression, we do a log transform, we do it both with and without the influential data point, but both times we learn that there is no linesar relationship. 


```{r}


xdml<-lm(Mean_speed~sex.preg, mdat3)
summary(xdml)
plot(fitted(xdml),residuals(xdml))
qqnorm(residuals(xdml))
dfbeta(xdml)
hist(residuals(xdml))

mdat3$logspeed<-log2(mdat3$Mean_speed)
log2<-lm(logspeed~sex.preg, mdat3)
summary(log2)
plot(fitted(log2),residuals(log2))
qqnorm(residuals(log2))
hist(residuals(log2))
dfbeta(log2)
#exponential trans
mdat3$expspeed<-exp(mdat3$Mean_speed)
xdml<-lm(expspeed~sex.preg, mdat3)
summary(xdml)
plot(fitted(xdml),residuals(xdml))
qqnorm(residuals(xdml))
dfbeta(xdml)
hist(residuals(xdml))
# now try the ANOVA

one.way <- aov(Mean_speed ~ sex.preg, data = mdat3)
summary(one.way)
par(mfrow=c(2,2))
plot(one.way)
par(mfrow=c(1,1))

 ggplot(mdat3, aes(x = sex.preg, y = Mean_speed)) +
  geom_point(cex = 1.5, pch = 1.0,position = position_jitter(w = 0.1, h = 0))
 
 xdml<-lm(Mean_speed~Sex, mdat3)
summary(xdml)
plot(fitted(xdml),residuals(xdml))
qqnorm(residuals(xdml))
dfbeta(xdml)
hist(residuals(xdml))

mdat3$logspeed<-log2(mdat3$Mean_speed)
log2<-lm(logspeed~Sex, mdat3)
summary(log2)
plot(fitted(log2),residuals(log2))
qqnorm(residuals(log2))
hist(residuals(log2))
dfbeta(log2)
#exponential trans
mdat3$expspeed<-exp(mdat3$Mean_speed)
xdml<-lm(expspeed~Sex, mdat3)
summary(xdml)
plot(fitted(xdml),residuals(xdml))
qqnorm(residuals(xdml))
dfbeta(xdml)
hist(residuals(xdml))
# now try the ANOVA

one.way <- aov(Mean_speed ~ Sex, data = mdat3)
summary(one.way)
par(mfrow=c(2,2))
plot(one.way)
par(mfrow=c(1,1))

 ggplot(mdat3, aes(x = Sex, y = Mean_speed)) +
  geom_point(cex = 1.5, pch = 1.0,position = position_jitter(w = 0.1, h = 0))
 welch.test(Mean_speed~Sex, data=mdat3)
 welch.test(Mean_speed ~ sex.preg, data = mdat3)
(Mean_speed ~ sex.preg, data = mdat3)
 welch_anova_test(Mean_speed ~ sex.preg, data = mdat3)
```
