---
title: "Three whales"
output: html_document
---
We are going to take a look at the speeds of the four whales that went all the way (and then also the energetics)


```{r}
library(tidyverse)
library(argosfilter)
library(stringr)
library(ggplot2)
library(ggmap)
library(lubridate)
```
Steps:

compare if the speeds are relatively stable
1) compare the instantaneous speed with scatter point graph
2) weekly speeds (or avg daily)

first we are going to graph these three
```{r}

speedth<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/summary stats/data compilations/speeds_all.csv")
avgsp3<-speedth%>%filter(Ptt%in%c("112699","121210","166123"))

register_google(key="<insertgooglekeyhere>")
pp<-c(mean(avgsp3$Longitude.x),-40)

m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=avgsp3, aes(x=Longitude.x, y=Latitude.x,col=as.factor(Ptt)),size=.5)

#so now that we have graphed it, looks like we will need to chop off some of the 166123
na166123<-avgsp3%>%filter(Ptt=="166123")
na166123$Date<-ymd(na166123$Date)
na166123<-na166123%>%filter(Date>=as.Date("2017-06-14"))
na112699<-avgsp3%>%filter(Ptt=="112699")
m121210<-avgsp3%>%filter(Ptt=="121210")

avgsp<-rbind(na166123,na112699,m121210)

#now lets graph again and see

pp<-c(mean(avgsp$Longitude.x),-40)

m <- get_map(location=pp,source="google",zoom=3,scale=2,maptype="satellite",color = "bw")
ggmap(m)+geom_path(data=avgsp, aes(x=Longitude.x, y=Latitude.x,col=as.factor(Ptt)),size=.5)
#looks good

```

Then we are doing the weekly speed graphs
```{r}
wkspeed<-read.csv("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/summary stats/data compilations/weeks_all.csv")

#grab the three
wkspthree<-wkspeed%>%filter(Ptt%in%c("112699","121210","166123"))

#set Ptt as factor
wkspthree$Ptt<-as.factor(wkspthree$Ptt)
#now graph them
ggplot(data=wkspthree)+
  geom_path(mapping=aes(x=week_num,y=Speed_km_h, color=Ptt), size=1)
#ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/summary stats/graphs/Three_speeds_week.png", height = 20, width = 20)
```

Now let's look at graphing the avg speeds



Now we are doing the avg speed graphs
```{r}
#graph julian day by speed
#but first we should probably get rid of some of the ridiculous speeds
avgspeed<-avgsp%>%filter(Speed_km_h<=22)
#have to add Julian day if I want them all on the same graph
avgspeed$julian<-yday(avgspeed$Date)
avgspeed$Ptt<-as.factor(avgspeed$Ptt)
ggplot(data=avgspeed)+
  geom_point(mapping=aes(x=julian,y=Speed_km_h, color=Ptt))

#ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/summary stats/graphs/Three_speeds_all_point.png", height = 20, width = 20)

#Now do it as the smooth
ggplot(data=avgspeed)+
  geom_smooth(mapping=aes(x=julian,y=Speed_km_h, color=Ptt))
#ggsave("/Users/michellemodest/Google Drive/Documents/UCSC/migration/R scripts/summary stats/graphs/Three_speeds_all_smooth.png", height = 20, width = 20)
#all of them seem to have decreased speeds by end of migration
ggplot(data=avgspeed)+
  geom_line(mapping=aes(x=julian,y=Speed_km_h, color=Ptt))
```



